Ground state energy of quantum dots using the coupled cluster method

### 2018-04-16 project_2/article/article.tex
Remove assumption on the SPF

\section{Theory}
    In this project we will study a system of $N$ interacting electrons. We will
    be looking at a Hamiltonian consisting of a one-body and a two-body part.
    The one-body part is given by
    \begin{align}
        h(\vf{r}_i)
        &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
    \end{align}
    where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
    The two-body part is the Coulomb interaction potential.
    \begin{align}
        v(\vf{r}_i, \vf{r}_j)
        &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
    \end{align}
    We thus get the total Hamiltonian
    \begin{align}
        H &= \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N v(\vf{r}_i, \vf{r}_j).
    \end{align}
    Next we define the reference Slater determinant as
    \begin{align}
        \ket{\Phi_0} &\equiv \ket{1, 2, \dots, N},
    \end{align}
    where $\ket{p}$ are the single particle functions that satisfies the
    equation
    \begin{align}
        h\ket{p} = \varepsilon_p\ket{p}.
    \end{align}
    They are the eigenfunctions to the one-body part of the Hamiltonian and are
    thus harmonic oscillator functions. Note that $\ket{\Phi_0}$ consists of the
    $N$ first occupied single particle functions. We will denote the occupied
    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the virtual states
    with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices with
    $p, q, r, s, \dots \in \{1, \dots, L\}$.

    \subsection{Second quantization}
        We now move over to a use second quantization. Employing the creation
        operators, $\acr{p}$, and the destruction operators, $\ade{p}$, we can
        write the Hamiltonian as
        \begin{align}
            H
            &=
            \sum_{pq}h_{q}^{p}\acr{p}\ade{q}
            + \sum_{pqrs}w_{rs}^{pq}\acr{p}\acr{q}\ade{s}\ade{r},
        \end{align}
        where the sums are general indices over all $L$ basis states and the
        matrix elements are defined as
        \begin{gather}
            h^{p}_{q} \equiv \bra{p}h\ket{q}, \\
            w_{rs}^{pq} \equiv \bra{pq}v\ket{rs}.
        \end{gather}
        In the case of the eigenstates we have that the one-body operator
        reduces to a diagonal matrix.
        \begin{align}
            h^{p}_{q}
            &= \bra{p}h\ket{q}
            = \varepsilon_q\braket{p}{q}
            = \varepsilon_q\delta_{pq},
        \end{align}
        where $\delta_{pq}$ is the Kronecker-Delta.

    \subsection{The coupled cluster approximation}
        We approximate the true wavefunction, $\ket{\Psi}$, of the system by the
        coupled cluster wavefunction, $\ket{\Psi_{\text{CC}}}$, defined as
        \begin{align}
            \ket{\Psi_{\text{CC}}}
            &\equiv e^{T}\ket{\Phi_0}
            = \para{
                \sum_{i = 0}^n
                \frac{1}{n!}T^n
            }\ket{\Phi_0},
        \end{align}
        where the \emph{cluster operator}, $T$, is given by a sum of
        $p$-excitation operators
        \begin{align}
            T &= T_1 + T_2 + \dots \\
            &=
            \sum_{ia}t_i^a\acr{a}\ade{i}
            + \para{\frac{1}{2!}}^2\sum_{ijab}
            t_{ij}^{ab}\acr{a}\acr{b}\ade{i}\ade{j}
            + \dots.
        \end{align}
        We will be looking at the \emph{doubles} approximation, that is
        \begin{align}
            T = T_{\text{CCD}} \equiv T_2.
        \end{align}

@@

        \subsubsection{Normal ordered Hamiltonian}
            To get to the coupled cluster equations we need to write the
            Hamiltonian on a normal ordered form. We use Wick's theorem on the
            Fermi vacuum\footnote{Fermi vacuum corresponds to the reference
            Slater.}. The one-body part of the Hamiltonian thus becomes
            \begin{align}
                h &= \sum_{pq}h^p_q\acr{p}\ade{q}
                = \sum_{pq}h^p_q\para{
                    \{\acr{p}\ade{q}\}
                    + \{
                        \wick{\c a_{p}^{\dagger} \c a_{q}}
                    \}
                }
                \\
                &= h_N + \sum_{pq}\varepsilon_i.
            \end{align}
            The two-body part yields
            \begin{align}
                v &= \sum_{pqrs}w^{pq}_{rs}\acr{p}\acr{q}\ade{s}\ade{r}
            \end{align}
            % TODO: Fill in the remainder. Preferably in the appendix.
            We can then write the Hamiltonian in terms of the reference energy
            and the normal ordered Hamiltonian.
            \begin{align}
                H = H_N + \bslat H\kslat.
            \end{align}
            The energy equation thus becomes
            \begin{align}
                E &= \bslat\bar{H}\kslat
                = E_{\text{ref}} + \bslat e^{-T}H_N e^T\kslat.
            \end{align}

>>>>

\section{Theory}
    In this project we will study a system of $N$ interacting electrons. We will
    be looking at a Hamiltonian consisting of a one-body and a two-body part.
    The one-body part is given by
    \begin{align}
        h(\vf{r}_i)
        &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
    \end{align}
    where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
    The two-body part is the Coulomb interaction potential.
    \begin{align}
        v(\vf{r}_i, \vf{r}_j)
        &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
    \end{align}
    We thus get the total Hamiltonian
    \begin{align}
        H &= \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N v(\vf{r}_i, \vf{r}_j).
    \end{align}
    Working in a basis of $L$ single particle functions, $\{\ket{p}\}_{p =
    1}^L$. We define the reference Slater determinant as
    \begin{align}
        \ket{\Phi_0} &\equiv \ket{1, 2, \dots, N},
    \end{align}
    i.e., a tensorproduct of the $N$ first single particle functions, $\ket{i}$,
    of the system. We call these single particle functions \emph{occupied} as
    they are contained in the Slater determinant.  We will denote the occupied
    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the \emph{virtual}
    states with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices
    with $p, q, r, s, \dots \in \{1, \dots, L\}$. In terms of sets of basis
    functions we can write this as
    \begin{align}
        \{\ket{p}\}_{p = 1}^L
        = \{\ket{i}\}_{i = 1}^N
        \cup
        \{\ket{a}\}_{a = N + 1}^L,
    \end{align}
    i.e., the general indexed states consists of both occupied and virtual
    states.

    \subsection{Second quantization}
        Employing the creation operators, $\acr{p}$, and the destruction
        operators, $\ade{p}$, we can write the Hamiltonian as
        \begin{align}
            H
            &=
            \sum_{pq}h_{q}^{p}\acr{p}\ade{q}
            + \sum_{pqrs}w_{rs}^{pq}\acr{p}\acr{q}\ade{s}\ade{r},
        \end{align}
        where the sums are general indices over all $L$ basis states and the
        matrix elements are defined as
        \begin{gather}
            h^{p}_{q} \equiv \bra{p}h\ket{q}, \\
            w_{rs}^{pq} \equiv \bra{pq}v\ket{rs}.
        \end{gather}
        Note that the two-body matrix elements are not antisymmetric yet.

    \subsection{The coupled cluster approximation}
        We approximate the true wavefunction, $\ket{\Psi}$, of the system by the
        coupled cluster wavefunction, $\ket{\Psi_{\text{CC}}}$, defined as
        \begin{align}
            \ket{\Psi_{\text{CC}}}
            &\equiv e^{T}\ket{\Phi_0}
            = \para{
                \sum_{i = 0}^n
                \frac{1}{n!}T^n
            }\ket{\Phi_0},
        \end{align}
        where the \emph{cluster operator}, $T$, is given by a sum of
        $p$-excitation operators
        \begin{align}
            T &= T_1 + T_2 + \dots + T_p \\
            &=
            \sum_{ia}t_i^a\acr{a}\ade{i}
            + \para{\frac{1}{2!}}^2\sum_{ijab}
            t_{ij}^{ab}\acr{a}\acr{b}\ade{i}\ade{j}
            + \dots,
        \end{align}
        where $t_{i\dots}^{a\dots}$
        We will be looking at the \emph{doubles} approximation, that is
        \begin{align}
            T = T_{\text{CCD}} \equiv T_2.
        \end{align}

@@

        \subsubsection{Normal ordered Hamiltonian}
            To get to the coupled cluster equations we need to write the
            Hamiltonian on a normal ordered form. We use Wick's theorem on the
            Fermi vacuum\footnote{Fermi vacuum corresponds to the reference
            Slater.}. The one-body part of the Hamiltonian thus becomes
            \begin{align}
                h &= \sum_{pq}h^p_q\acr{p}\ade{q}
                = \sum_{pq}h^p_q\para{
                    \{\acr{p}\ade{q}\}
                    + \{
                        \wick{\c a_{p}^{\dagger} \c a_{q}}
                    \}
                }
                \\
                &= h_N + \sum_{pq}\varepsilon_i.
            \end{align}
            The two-body part yields
            \begin{align}
                v &= \sum_{pqrs}w^{pq}_{rs}\acr{p}\acr{q}\ade{s}\ade{r}
            \end{align}
            % TODO: Fill in the remainder. Preferably in the appendix.
            We can then write the Hamiltonian in terms of the reference energy
            and the normal ordered Hamiltonian.
            \begin{align}
                H = H_N + \bslat H\kslat.
            \end{align}
            The energy equation thus becomes
            \begin{align}
                E &= \bslat\bar{H}\kslat
                = E_{\text{ref}} + \bslat e^{-T}H_N e^T\kslat.
            \end{align}
            In this equation the unknowns are the cluster amplitudes
